---
title: "testing hurdle_cumulative family"
author: "Stephen Wild"
date: "14/01/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Surveys often contain ordinal variables, like Likert scales. These can be tough to deal with. One great thing about brms is its ability to deal with ordinal data using a cumulative logit (or probit, or cauchit, or...) model, rather than treating the data as continuous.

One issue that often arises is how to model responses when respondents answer "Don't know." This happens often. Many times you can only respond to a Likert scale question if you know something about the topic! 

In this way, we can treat the responses to the Likert scale as having to overcome a hurdle: You need to know something about the topic to have an opinion on it.

Thankfully, brms let us model this type of data. Let's start by loading the packages we'll need.


```{r load packages, results = 'hide'}
devtools::load_all("/Users/stephenwild/Desktop/Stats_stuff/brms")
library(tidyverse)

```

To make use our hurdle ordinal model, we'll create some test data. 
```{r test data}

# create test data. We'll compare our results to these values
N <- 2500
P <- 5
cutpoints <- c(1, -1)
y <- rep(0, N)
betas <- c(-.15, -.5, 3, .5, -1)
gammas <- c(3, -2, -2, -1, -.5)


# Build predictors and response
set.seed(143140914) # set seed for reproducibility
X <- matrix(runif(N * P), N, P)
p2 <- plogis(cutpoints[1] + X %*% betas)
p3 <- plogis(cutpoints[2] + X %*% betas)


# create thresholds
p1 <- 1 - p2
p2 <- p2 - p3


# sample based on thresholds
for(i in 1:N){
  y[i] <- sample(1:3, 
                 size = 1,
                 prob = c(p1[i], p2[i], p3[i]))
}


# Simulate "Don't knows"
# It must be 0 to serve as the hurdle
p <- plogis(X %*% gammas)
dk <- rbinom(N, 1, p)
y[dk == 1] = 0 # replaces "true' values respondents would give if they knew about topic

# put it all together in a dataframe
df <- data.frame(y = y, X)

```

And with that, we have our data. Let's compare it and see what it looks like.

```{r plot data}
ggplot(data = df) +
  geom_bar(mapping = aes(x = y),
           fill = "firebrick") +
  labs(title = "Response values in simulated data") +
  theme_bw()

```

Perfect. with that we can model our data. We'll see if the values we get are similar to the parameters we simulated.

```{r run model, results = 'hide'}
mod <- brm(bf(y ~ X1 + X2 + X3 + X4 + X5,
           hu ~ 0 + X1 + X2 + X3 + X4 + X5),
           family = hurdle_cumulative(),
           data = df,
           cores = 4,
           chains = 4)
```

```{r display results of mod}
mod

```

Success! our values are similar. That's good and what we want to see. Let's also try other links to make sure they work too.

```{r other links, , results = 'hide'}

mod2 <- brm(bf(y ~ X1 + X2 + X3 + X4 + X5,
              hu ~ 0 + X1 + X2 + X3 + X4 + X5),
           family = hurdle_cumulative(link = "probit"),
           data = df,
           cores = 4,
           chains = 4)
mod3 <- brm(bf(y ~ X1 + X2 + X3 + X4 + X5,
               hu ~ 0 + X1 + X2 + X3 + X4 + X5),
            family = hurdle_cumulative(link = "cauchit"),
            data = df,
            cores = 4,
            chains = 4)
```

```{r display results of other links}
mod2
mod3
```

Finally, we want to test the post-processing functions, like `loo_compare()` and `pp_check()`. This is important, because a lot of users use brms because a) Stan can be confusing, and b) it's easy to work with brmsfit objects.

```{r test post-processing functions}

loo_compare.brmsfit(add_criterion(mod, criterion = "loo"), 
                    add_criterion(mod2, criterion = "loo"),
                    add_criterion(mod3, criterion = "loo"))

pp_check(mod, type = "bars", ndraws = 50) +
  labs(title = "pp_check for test data") +
  theme_bw()

```


## test with merkel data used by Octavio Medina
Finally, we want to test it on real data. Thankfully, Octavio Medina has us covered. He has a great [blog post](https://octavio.me/posts/dont-knows/) on using Stan to model "Don't knows." He also produced a [custom brms family](https://discourse.mc-stan.org/t/modeling-non-response-in-ordinal-survey-data/25151/2) that can be used.

I'm going to skip a lot of what he did with Stan and the custom family and simply load his data.

```{r load and test merkel data, results = 'hide'}
merkel <- read.csv("https://raw.githubusercontent.com/octmedina/zi-ordinal/main/merkel_data.csv")

mod_merkel <- brm(bf(confid_merkel ~ edu + race + income + party,
                     #disc ~ 1 + edu,
                     hu ~ 1 + edu + race + income + party),
                  data = merkel,
                  family = hurdle_cumulative(),
                  cores = 4,
                  chains = 4)
```

```{r display merkel output}
mod_merkel
pp_check(mod_merkel, type ="bars", ndraws = 50) + 
  labs(title = "pp_check for Merkel data") +
  theme_bw()
```

And there is is. The hurdle_cumulative family works.
